---
title: "Univariate Viz"
---

Use this file for practice with the **univariate viz** in-class activity.  Refer to the class website for details.

```{r}
# Import data
hikes <- read.csv("https://mac-stat.github.io/data/high_peaks.csv")
```
## Excercise 1
- What features would we like a visualization of the categorical difficulty rating variable to capture?
- What about a visualization of the quantitative elevation variable?
```{r}
head(hikes)
```
##Exercise 2: Load tidyverse
```{r}
# Load the package
library(tidyverse)
```
##Exercise 3: Bar Chart of Ratings - Part 1
- How many hikes fall into each category?
- Are the hikes evenly distributed among these categories, or are some more common than others?
```{r}
ggplot(hikes, aes(x = rating))
```
Think about:

- What did this do? What do you observe?
- What, in general, is the first argument of the ggplot() function?
- What is the purpose of writing x = rating?
- What do you think aes stands for?!?
##Exercise 4: Bar Chart of Ratings - Part 2
``` {r}
# added geoem_bar() and bars appeared!
ggplot(hikes, aes(x = rating)) +
  geom_bar()
# added x=rating and y=number of hikes and it changed the labels of the x and y axes
ggplot(hikes, aes(x = rating)) +
  geom_bar() +
  labs(x = "Rating", y = "Number of hikes")
# added fill blue and low and behold it filled blue
ggplot(hikes, aes(x = rating)) +
  geom_bar(fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")
# added color=orange and it gave an orange outline
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue") +
  labs(x = "Rating", y = "Number of hikes")
# added theme_minimal and it mader the background white instead of grey
ggplot(hikes, aes(x = rating)) +
  geom_bar(color = "orange", fill = "blue")  +
  labs(x = "Rating", y = "Number of hikes") +
  theme_minimal()
```

## Exercise 5: Bar Chart Follow-up
### Part a

Reflect on the ggplot() code.

- What’s the purpose of the +? When do we use it? *to add a seperate command in the same line of code*
- We added the bars using geom_bar()? Why “geom”?
- What does labs() stand for?
- What’s the difference between color and fill? *fill "fills" the color in, while color creates an outline*

### Part b
In general, bar charts allow us to examine the following properties of a categorical variable:

- **observed categories**: What categories did we observe? *difficult, easy, and moderate*
- **variability between categories**: Are observations evenly spread out among the categories, or are some categories more common than others? *moderate is most common, then easy, then difficult*

We must then *translate* this information into the *context* of our analysis, here hikes in the Adirondacks. Summarize below what you learned from the bar chart, in context.

*most people (almost 30) go on moderate hikes, and the least amount of people (less than 10) go on difficult hikes*

### Part c

Is there anything you don’t like about this barplot? For example: check out the x-axis again.

*i dont like that the order is not easy to difficult.*

## Exercise 6: Sad Bar Chart

```{r}
ggplot(hikes, aes(x = elevation)) +
  geom_bar(color = "pink", fill = ("pink")) + 
  labs(x = "elevation", y = "Number of hikes") +
  theme_minimal()
```
