---
title: "Mulivariate Viz"
---

Use this file for practice with the **mulivariate viz** in-class activity.  Refer to the class website for details.

```{r}
library(ggplot2)
```

## The Story
``` {r}
# Import and check out data
education <- read.csv("https://mac-stat.github.io/data/sat.csv")
head(education)
```
## Exercise 1: SAT scores
### Part a
Construct a plot of how the average sat scores vary from state to state. (Just use 1 variable – sat not State!)

### Part b
Summarize your observations from the plot. Comment on the basics: range, typical outcomes, shape. (Any theories about what might explain this non-normal shape?)

## Exercise 2: SAT Scores vs Per Pupil Spending & SAT Scores vs Salaries
The first question we’d like to answer is: Can the variability in sat scores from state to state be partially explained by how much a state spends on education, specifically its per pupil spending (expend) and typical teacher salary?

part 1
```{r}
ggplot(education, aes(x=sat, y=expend)) + geom_smooth(method= "lm")
```
part 2

```{r}
ggplot(education, aes(x=sat, y=salary)) + geom_smooth(method= "lm")
```

## Exercise 3: SAT Scores vs Per Pupil Spending and Teacher Salaries
Construct one visualization of the relationship of sat with salary and expend. HINT: Start with just 2 variables and tweak that code to add the third variable. Try out a few things!

```{r}
ggplot(education, aes(x=sat, y=expend, 
                    color=salary)) + geom_point()
```

## Exercise 4: Another way to Incorporate Scale
It can be tough to distinguish color scales and size scales for quantitative variables. Another option is to discretize a quantitative variable, or basically cut it up into categories.

Construct the plot below. Check out the code and think about what’s happening here. What happens if you change “2” to “3”?

```{r}
ggplot(education, aes(x=sat)) + geom_density()
```


```{r}
ggplot(education, aes(y = sat, x = salary, color = cut(expend, 3))) + 
  geom_point() + 
  geom_smooth(se = FALSE, method = "lm")
```

## Exercise 5: Finally an Explanation
It’s strange that SAT scores seem to decrease with spending. But we’re leaving out an important variable from our analysis: the fraction of a state’s students that actually take the SAT. The fracCat variable indicates this fraction: low (under 15% take the SAT), medium (15-45% take the SAT), and high (at least 45% take the SAT).

```{r}
ggplot(education, aes(x=frac))
```

### Part a
Build a univariate viz of fracCat to better understand how many states fall into each category.

### Part b
Build 2 bivariate visualizations that demonstrate the relationship between sat and fracCat. What story does your graphic tell and why does this make contextual sense?

### Part c
Make a trivariate visualization that demonstrates the relationship of sat with expend AND fracCat. Highlight the differences in fracCat groups through color AND unique trend lines. What story does your graphic tell?
Does it still seem that SAT scores decrease as spending increases?

### Part d
Putting all of this together, explain this example of Simpson’s Paradox. That is, why did it appear that SAT scores decrease as spending increases even though the opposite is true?


